Type-Safe Modular Parsing

### 2017-01-12 paper/sections/Discussion.tex
Revised 5.1

Just as many parser generators do\huang{???}, our pattern also supports overriding part of existing parsers, but in a type-safe way.
We can easily update existing parsing rules with new implementations, or eliminate them in extended parsers. As an illustration, suppose we have an untyped lambda abstraction case in a base parser as below. \inlinecode{pLam} parses a lambda symbol, an identifier, a dot and an expression in sequence. Some irrelevant details are omitted.

\lstinputlisting[linerange=7-12]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM

@@

\lstinputlisting[linerange=16-22]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM

@@

It is natural to keep overriding existing parsers, whereas another potential use of this pattern is to reuse old versions of a parser. For instance,
we have a \lstinline{NewParser} that extends \lstinline{ExtParser}, with some more cases added, but for the lambda case we want to go back to the untyped one.
It can be done by copying the code from \lstinline{BaseParser} again, yet does not achieve code reuse. Instead, we can instantiate an instance of \lstinline{BaseParser} to obtain the old \lstinline{pLam}:
\haoyuan{I don't know if this works or makes sense.}

\lstinputlisting[linerange=26-32]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM

@@

\begin{figure}[t]
\begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
\lstinputlisting[linerange=44-59]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
&
\lstinputlisting[linerange=63-78]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
\end{tabular}
\caption{Language \lstinline{VarExpr} (left) and language \lstinline{TypedLam} (right).}\label{fig:lng-components}
\end{figure}

@@

\lstinputlisting[linerange=82-91]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
Such composition shows perfect modularity. The only
drawback is that the glue code of composition appears to be boilerplate. As shown above, we are combining ASTs, parsers and pretty printers of
\lstinline{VarExpr} and \lstinline{TypedLam} respectively. Such a pattern refers to \textit{family polymorphism}[]. Unfortunately Scala does not
support this feature, nonetheless, one can avoid such boilerplate by using metaprogramming techniques.

>>>>

As many syntactically extensible parsers, our approach also supports changing part of existing parsers, but in a type-safe way. We can easily update existing parsing rules with new implementations, or eliminate them in extended parsers. As an illustration, suppose we have an untyped lambda abstraction case in a base parser as below. \inlinecode{pLam} parses a lambda symbol, an identifier, a dot and an expression in sequence. Some irrelevant details are omitted.

\lstinputlisting[linerange=76-81]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=BASEPARSER_UNTYPEDLAM

@@

\lstinputlisting[linerange=85-90]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=EXTPARSER_TYPEDLAM

@@

It is natural to keep overriding existing parsers, whereas another potential use of this pattern is to reuse old versions of a parser. For instance,
we have a \lstinline{NewParser} that extends \lstinline{ExtParser}, with some more cases added, but for the lambda case we want to go back to the untyped one.
It can be achieved by instantiating an instance of \lstinline{BaseParser} to obtain the old \lstinline{pLam}. However, we must have \inlinecode{lazy} modifier for all the \inlinecode{pLam} starting from \inlinecode{BaseParser}.

\huang{i finally managed to make it work, with two overriding. i think this case may not be very common and the solution is kind of unly. not sure whether to keep this case or not}\haoyuan{I don't know if this works or makes sense.}

\lstinputlisting[linerange=94-102]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=NEWPARSER_UNTYPEDLAM

@@

\begin{figure}[t]
\begin{tabular}{m{0.42\linewidth}m{0.52\linewidth}}
\lstinputlisting[linerange=113-128]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VAREXPR
&
\lstinputlisting[linerange=132-147]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_TYPEDLAM
\end{tabular}
\caption{Language \lstinline{VarExpr} (left) and language \lstinline{TypedLam} (right).}\label{fig:lng-components}
\end{figure}

@@

\lstinputlisting[linerange=151-160]{../Scala/Parser/src/PaperCode/Sec5/Code1.scala}% APPLY:linerange=LANGUAGE_COMPONENTS_VARLAMEXPR
Such composition shows perfect modularity. The only
drawback is that the glue code of composition appears to be boilerplate. As shown above, we are combining ASTs, parsers and pretty printers of
\lstinline{VarExpr} and \lstinline{TypedLam} respectively. Such a pattern refers to \textit{family polymorphism}[]. Unfortunately Scala does not
support this feature, nonetheless, one can avoid such boilerplate by using metaprogramming techniques.
